---
title: "GSEA of Proteomics"
author: "Evangelyn Sim"
date: "2/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Introduction

Obtain Proteomics differential analysis outcome: YAP shRNA D28 Proteomics results_ES.txt



### Used libraries and functions

* java/1.8.0_66
* parallel
```{r}
library(ggplot2)
library(grid)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(dplyr)

```



### Read files 
```{r}

name = read.delim("/group/card2/Evangelyn_Sim/Collaboration_Kev_UoM/Sequencing_ATAC_RNA/20190531_RNA_run1/seqaln/rename/mrna_fulllen_pe_strrev.mx.chr", header = T, sep = "\t")
name$Gene.names = gsub(".*_","",name$Geneid)
name = name[,c(1,7)]

prote = read.delim("/group/card2/Evangelyn_Sim/Collaboration_Kev_UoM/Sequencing_ATAC_RNA/20190530_ATAC_run1/R/1.1kbpTSS/8.intersectwithproteomic/YAP shRNA D28 Proteomics results_ES.txt", header = T, sep = "\t")

mg = merge(prote, name, by = "Gene.names")
mg = mg[,c(21,2:20)]

write.table(mg, file = "/group/card2/Evangelyn_Sim/Collaboration_Kev_UoM/Sequencing_ATAC_RNA/Github/2021_UoM_Yap_shRNA_nuclei_RNAseq_ATACseq/output/YAP_shRNA_D28_Proteomics_results.xls", col.names = T, row.names = F, sep = "\t")


```


### Rank Genes and change mouse gene name to human homolog

Following generated edgeR spreadsheet, use the logFC and p.Value to generate a rank score using the following scripts.

rnkgenM2H.sh

      #!/bin/bash

      #Specify the input file
      XLS=$1
      #Specify the gene ID column
      ID=$2
      #Specify the fold change value column
      FC=$3
      #Specify the raw p-value column
      P=$4
      #Specify ortholog maping
      ORTH=$5


      RNK=${XLS}.rnk
      HUM=${RNK}.hum.rnk

      sed 1d $XLS | tr -d '"' \
      | awk -v I=$ID -v F=$FC -v P=$P '{FS="\t"} {print $I, $F, $P}' \
      | awk '$2!="NA" && $3!="NA"' \
      | awk '{s=1} $2<0{s=-1} {print $1"\t"s*-1*log($3)/log(10)}' \
      | sort -k2gr | sed 's/inf$/330/'> $RNK

      sed 's/_/\t/' $RNK \
      | sort -k 1b,1 \
      | join -1 2 -2 1 $ORTH - \
      | awk '{OFS="\t"} {print $0,$5*$5}' \
      | sort -k6gr \
      | awk '!arr[$4]++' \
      | awk '{OFS="\t"} !arr[$3]++ {print $3,$5}' \
      | sort -k2gr > $HUM




Run rnkgenM2H.sh to generate .rnk files

      #!/bin/bash
      for XLS in *xls ; do
         ./rnkgen.sh $XLS 1 5 6 mouse2human.txt.sort 1 2 3 ;
      done


### Run GSEA using ranked genes

Subject the generated .rnk files along with .gmt file sand run the following scripts to perform gene set enrichment analysis. 

Download gmt files from GSEA webpage
http://www.gsea-msigdb.org/gsea/login.jsp;jsessionid=C4D3892651A8792A331D7B32E9D2269C

rungsea.sh

      #!/bin/bash

      run_gsea(){
      RNK=$1
      GMT=$2
      echo /group/card2/Evangelyn_Sim/NGS/app/gsea-3.0.jar $RNK $GMT
      java -Xmx4096m -cp /group/card2/Evangelyn_Sim/NGS/app/gsea-3.0.jar xtools.gsea.GseaPreranked  \
      -gmx $GMT -collapse false -mode Max_probe \
      -norm meandiv -nperm 1000 -rnk $RNK -scoring_scheme classic \
      -rpt_label ${RNK}.${GMT} -include_only_symbols true -make_sets true \
      -plot_top_x 20 -rnd_seed timestamp -set_max 5000 -set_min 10 -zip_report false \
      -out . -gui false
      }
      export -f run_gsea

      parallel -j5 run_gsea ::: *rnk ::: *gmt


### Parse GSEA outcome

```{bash}
#!/bin/bash
echo 'GeneSetName	GeneSetSize	ES	NES	p-val	FDR	FWER' > header.txt

for GSEADIR in `ls | grep GseaPreranked | grep -v xls$` ; do
  awk '{FS="\t"} {OFS="\t"} $8<0.05 {print $1,$4,$5,$6,$7,$8,$9} ' $GSEADIR/gsea_report_for_na_*xls \
  | cat header.txt - > $GSEADIR.xls
done



```


### Plot GSEA result

```{r}
files = list.files(path = "/group/card2/Evangelyn_Sim/Collaboration_Kev_UoM/Sequencing_ATAC_RNA/20190530_ATAC_run1/R/1.1kbpTSS/10.proteomic_gsea", pattern = ".*go.xls$", full.names = T)
mx = lapply(files, read.delim, header=T)


for(l in 1:length(mx)){
  
  mx[[l]]$GeneSetName = gsub("GO_", "", mx[[l]]$GeneSetName)
  
  mxRU= mx[[l]]
  mxRU= mxRU[order(mxRU$NES, decreasing = T), ]
  mxRU= mxRU[c(1:5),]
  mxRU$colour = "blue"
  
  mxRD= mx[[l]]
  mxRD= mxRD[order(mxRD$NES), ]
  mxRD= mxRD[c(1:5),]
  mxRD$colour = "red"
  
  ES_all = rbind(mxRU, mxRD)
  
}
```


```{r, fig.height=6, fig.width= 15}

par(las =2)
par(mar=c(3,55,5,2))
  
fig = barplot(rev(ES_all$NES), 
                horiz = T, 
                col = ES_all$colour, 
                names.arg = rev(ES_all$GeneSetName) , 
                cex.axis = 1.5, cex.names = 1.25)


```
